---
title: "Breakpoint methods in R"
author: "Emily Hardison and Krista Kraskura"
date: "1/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Good initial resource [here](https://lindeloev.github.io/mcp/articles/packages.html)

### What are methods used by researchers in the field:

1. "Piecewise linear regression models **(Satistica 13)**"
    - https://www.sciencedirect.com/science/article/pii/S0306456519304814
2. Matt Gilbert used R Studio (R Core Team, 2014) except for the segmented regression analysis which, along with all data presentation, was done using **Prism v.8.3.** 
    - https://academic.oup.com/conphys/article/8/1/coaa036/5823905#202778356
3. Katja Anttila was "finding best fits for two linear regression lines and calculating their
intercept (**Sigmaplot 11.0**, Systat Software Inc., www.sigmaplot.com)."
    - https://www.nature.com/articles/ncomms5252
  
  
### What are methods available in R that we will compare below:

1. package `segmented`
    - https://rpubs.com/MarkusLoew/12164
2. package `EnvCpt`
    - Detection of Structural Changes in Climate and Environment Time Series
3. package `strucchange`
    - function: `breakpoints()`
    - function: `efp()`
    - funciton: `Fstats()`
4. 


#### Install packages to be used:
```{r}
# packages for breakpoint analysis
library(mcp)
library(EnvCpt) 
library(segmented)
library(strucchange) # Fstats 

# additional packages
library(tidyverse)
library(ggplot2)

```


```{r}
set.seed(423485)
```


#### Viewving the original dataset


```{r}
data<-read.csv("../data/abt_example_data.csv")
data.max<-data %>% 
  dplyr::group_by(channel) %>% 
  summarise(max_bpm = max(bpm))

data_full<-left_join(data, data.max)

# str(data_full)
data_full2 <- data_full %>% 
filter(
    bpm == max_bpm) %>% 
  rename(tpeak=temp_mean)

data_full3<-left_join(data_full, data_full2, by = "channel") %>% 
  filter(temp_mean<=tpeak) %>% 
  mutate(
    kelvin = temp_mean + 273.15,
    arrhenius_temp = (10^3)/(kelvin)) %>% 
  as.data.frame()
data_full3<-data_full3[, c("channel", "timepoint.x", "bpm.x", "temp_mean", "arrhenius_temp", "tpeak")]

ggplot(data_full3, aes(1000/celsius.to.kelvin(temp_mean), log(bpm.x)))+
  geom_point()+
  theme_bw()+
  facet_grid(.~channel)+
  labs(
        x = expression(paste(Arrhenius~temperature~(10^3*K^-1))),
        y = expression(paste(ln(f[Hmax]))))
  
list.data<-split(data_full3,  data_full3$channel)
ch1<-as.data.frame(list.data[1]) 
colnames(ch1)<-colnames(data_full3)
ch2<-as.data.frame(list.data[2])
colnames(ch1)<-colnames(data_full3)
ch6<-as.data.frame(list.data[3])

```

```{r}

library(mcp)

# Plot it
plot(df)
abline(v = c(30, 70), col="red")


```



#### Segmented 

* uses bootstrap 
* can set seed
* can provide primer values
* obtain SE around estimated parameters

```{r}
### Fit a lm to each fish ID

df<- data_full3[data_full3$channel=="Ch1",]

for (i in 1:300){
  if(i == 100 |i == 200 |i == 300){
    set.seed(i)
  }
      
    fish.lm <- lm(log10(bpm.x) ~ arrhenius_temp, data = df)
    ### Run a breakpoint test on all the data (since you need to estimate the breakpoints, I split the files up by temperature)
    
    my.seg <- segmented(fish.lm,
                      seg.Z = ~ arrhenius_temp,
                      psi = list(arrhenius_temp = c(3.44)))
    
    my.fitted <- fitted(my.seg)
    my.model <- data.frame(a.temp = df$arrhenius_temp, bpm = my.fitted)

    p <- ggplot(df, aes(x = arrhenius_temp, y = log10(bpm.x))) + geom_line()
    plot.data <- p + geom_line(data = my.model, aes(x = a.temp, y = bpm), colour = "red")+
      geom_vline(aes(xintercept = my.seg$psi[2]))+
      labs(
        x = expression(paste(Arrhenius~temperature~(10^3*K^-1))),
        y = expression(paste(ln(f[Hmax]))))

    print(c(i, my.seg$psi[2]))
  
}

    




```


#### SiZer

* uses bootstrap 
* can set seed
* can provide primer values
* obtain CI around estimated parameters, can specify sign level


```{r}

```


```{r}
library(strucchange)

strucchange::breakpoints



```






